ToothGrowth data analysis
=========================

**Load the ToothGrowth data and perform some basic exploratory data analysis**

The below data plot shows that average tooth length has a linear relationship with amount of dose. Higher the dose, longer the tooth.
```{r echo=TRUE}
library(ggplot2)
lenBYdose <- NULL
data(ToothGrowth)
lenBYdose <- aggregate(ToothGrowth$len, by=list(ToothGrowth$dose), FUN=mean)
names(lenBYdose) <- c("len", "dose")

ggplot(lenBYdose, aes(dose, len, label=round(dose,1))) + 
  geom_line() +
  ylab("Average tooth length") +
  xlab("Dose in milligrams") +
  ggtitle("Average Tooth Length by Dose") +
  geom_text()

```

The following plot is a different view of the positive relationship between tooth length and the dose levels of
Vitamin C for both delivery methods. The difference in tooth length is significant when dose amount is 0.5mm or 1mm. At 2mm, the difference becomes less significant.

```{r echo=TRUE}
ggplot(ToothGrowth, aes(as.factor(dose), len, fill=supp, label=round(len,1))) +
  geom_bar(stat="identity") + 
  facet_grid(. ~ supp) +
  xlab("Dose in milligrams") +
  ylab("Tooth length") +
  guides(fill=guide_legend(title="Supplement type"))

```

**Basic summary**
```{r echo=TRUE}
summary(ToothGrowth)

```
ToothGrowth is a dataset of type `r class(ToothGrowth)` with:

number of observations: `r nrow(ToothGrowth)`

number of columns: `r ncol(ToothGrowth)`

column names: `r names(ToothGrowth)`

**Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. State your conclusions**

We will us use T-Tests to compare tooth growth by supplement type and dosage amount.

*First, we will compare tooth growth by supplement type*
```{r echo=TRUE}
t.test(len ~ supp, data=ToothGrowth)

```
The p-value of `r round(as.numeric(t.test(len ~ supp, data=ToothGrowth)[3]), 2)` indicates that if we were to repeat our experiment many, many times (each time we select `r nrow(ToothGrowth)/2` observations for each of supplement type at random and compute the sample mean) then `r round(as.numeric(t.test(len ~ supp, data=ToothGrowth)[3]), 2)*100` times out of 100 we can expect to see a sample mean greater than mean of the group for each supplement type.

The 95% confidence interval shows that there is a 95% chance that the confidence interval `r t.test(len ~ supp, data=ToothGrowth)[[4]][1:2]` contains the true population mean. If repeated samples were taken and the 95% confidence interval computed for each sample, 95% of the intervals would contain the population mean.


*Next, we will compare tooth growth by dosage amount by subsetting data in 3 pairs of doses.*
```{r echo=TRUE}
t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(0.5, 1.0)))
```
The p-value of `r t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(0.5, 1.0)))[3]` indicates that if we were to repeat our experiment many, many times (each time we select `r nrow(ToothGrowth)/2` observations for each of supplement type at random and compute the sample mean) then `r round(as.numeric(t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(0.5, 1.0)))[3]), 4)*100` times out of 100 we can expect to see a sample mean greater than mean of the group for each supplement type.

The 95% confidence interval shows that there is a 95% chance that the confidence interval `r t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(0.5, 1.0)))[[4]][1:2]` contains the true population mean. If repeated samples were taken and the 95% confidence interval computed for each sample, 95% of the intervals would contain the population mean.



```{r echo=TRUE}
t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(1.0, 2.0)))
```
The p-value of `r t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(1.0, 2.0)))[3]` indicates that if we were to repeat our experiment many, many times (each time we select `r nrow(ToothGrowth)/2` observations for each of supplement type at random and compute the sample mean) then `r round(as.numeric(t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(1.0, 2.0)))[3]), 4)*100` times out of 100 we can expect to see a sample mean greater than mean of the group for each supplement type.

The 95% confidence interval shows that there is a 95% chance that the confidence interval `r t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(1.0, 2.0)))[[4]][1:2]` contains the true population mean. If repeated samples were taken and the 95% confidence interval computed for each sample, 95% of the intervals would contain the population mean.

```{r echo=TRUE}
t.test(len ~ dose, data=subset(ToothGrowth, dose==c(0.5, 2.0)))
```
The p-value of `r t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(0.5, 2.0)))[3]` indicates that if we were to repeat our experiment many, many times (each time we select `r nrow(ToothGrowth)/2` observations for each of supplement type at random and compute the sample mean) then `r round(as.numeric(t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(0.5, 2.0)))[3]), 4)*100` times out of 100 we can expect to see a sample mean greater than mean of the group for each supplement type.

The 95% confidence interval shows that there is a 95% chance that the confidence interval `r t.test(len ~ dose, data=subset(ToothGrowth, dose %in% c(0.5, 2.0)))[[4]][1:2]` contains the true population mean. If repeated samples were taken and the 95% confidence interval computed for each sample, 95% of the intervals would contain the population mean.


**State the assumptions needed for your conclusions**

1. Samples are random, independent, and come from normally distributed population with unknown but equal variances. 
2. Populations follow a Gaussian distribution.

end of document
===============
