ToothGrowth data analysis
=========================
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 
Load the ToothGrowth data and perform some basic exploratory data analyses 
Provide a basic summary of the data.
Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
State your conclusions and the assumptions needed for your conclusions. 
Some criteria that you will be evaluated on
Did you  perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
Did the student perform some relevant confidence intervals and/or tests?
Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 
Did the student describe the assumptions needed for their conclusions?

**Load the ToothGrowth data and perform some basic exploratory data analysis**

The below data plot shows that average tooth length has a linear relationship with amount of dose. Higher the dose, longer the tooth.
```{r echo=TRUE}
library(ggplot2)
lenBYdose <- NULL
data(ToothGrowth)
lenBYdose <- aggregate(ToothGrowth$len, by=list(ToothGrowth$dose), FUN=mean)
names(lenBYdose) <- c("len", "dose")

ggplot(lenBYdose, aes(dose, len, label=round(dose,1))) + 
  geom_line() +
  ylab("Average tooth length") +
  xlab("Dose in milligrams") +
  ggtitle("Average Tooth Length by Dose") +
  geom_text()

```

The following plot is a different view of the positive relationship between tooth length and the dose levels of
Vitamin C for both delivery methods. The difference in tooth length is significant when dose amount is 0.5mm or 1mm. At 2mm, the difference becomes less visible.

```{r echo=TRUE}

ggplot(ToothGrowth, aes(as.factor(dose), len, fill=supp, label=round(len,1))) +
  geom_bar(stat="identity") + 
  facet_grid(. ~ supp) +
  xlab("Dose in milligrams") +
  ylab("Tooth length") +
  guides(fill=guide_legend(title="Supplement type"))

```



**Basic summary**
```{r echo=TRUE}
summary(ToothGrowth)
```
ToothGrowth is a dataset of type `r class(ToothGrowth)` with:

number of observations: `r nrow(ToothGrowth)`

number of columns: `r ncol(ToothGrowth)`

column names: `r names(ToothGrowth)`


**Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. State your conclusions**

Let's use T-Tests to compare tooth growth by supplement type and dosage amount.

*First, we will compare tooth growth by supplement type*
```{r echo=TRUE}
t.test(len ~ supp, data=ToothGrowth)
```

*Next, we will compare tooth growth by dosage amount*
```{r echo=TRUE}
t.test(len ~ dose, data=subset(ToothGrowth, dose==c(0.5, 1.0)))
```


**State the assumptions needed for your conclusions**
1. Samples are random, independent, and come from normally distributed population with unknown but equal variances. 
2. Populations follow a Gaussian distribution.
